---
title: "Getting started with LLMRcontent"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with LLMRcontent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

LLMRcontent treats model labels as measured variables. A study defines the
instrument, develops candidate protocols on designated rows, validates one
locked protocol against held-out human labels, and carries the remaining
classification error into corpus estimates. Robustness checks show how
specified measurement choices affect the estimand, while replication archives
connect results to recorded calls.

```{r}
library(LLMRcontent)
```

## Coding with held-out validation

The codebook states the construct and its category boundaries.

```{r}
cb <- codebook(
  name = "policy appraisal",
  unit = "one statement",
  categories = list(
    cb_category("supportive", "Presents the proposal as beneficial."),
    cb_category("critical", "Presents the proposal as harmful or deficient.")
  ),
  version = "1.0"
)

cb
```

The gold set stores its split assignment and seals the test split. Stable unit
identifiers provide the later link between audited rows and the coded corpus.
Before collecting labels, `gold_size()` can compare candidate sample sizes for
a target agreement-interval width. The development split supports protocol
selection; the test split is reserved for evaluating the selected protocol.

```{r}
gold_data <- data.frame(
  id = sprintf("g%02d", seq_len(40)),
  text = c(
    paste("clear public benefit case", seq_len(20)),
    paste("serious public harm case", seq_len(20))
  ),
  label = rep(c("supportive", "critical"), each = 20),
  stringsAsFactors = FALSE
)

set.seed(110)
gold <- gold_set(
  gold_data,
  text = "text",
  label = "label",
  id = "id",
  split = c(dev = 0.5, test = 0.5)
)

gold
```

A locked protocol hashes the codebook, prompt, model configuration, and parser.
In a full study, `tune_protocol()` compares candidates on the development
split before one is locked; this compact example fixes one protocol so the
holdout result and corrected prevalence remain the focus. The same locked
object must be used for validation and corpus coding.

```{r}
cfg <- LLMR::llm_config("groq", "offline-demo", temperature = 0)
locked <- protocol_lock(
  protocol(cb, cfg, label = "keyword protocol")
)

locked
```

Execution functions accept a `.runner` through an interface in which a function
receives request rows and returns response rows; this offline response function
applies a fixed keyword rule without contacting a provider.

```{r}
offline_coder <- function(requests, ...) {
  text <- tolower(requests$text)
  requests$response_text <- ifelse(
    grepl("benefit", text, fixed = TRUE), "supportive", "critical"
  )
  requests$response_text[grepl("benefit case (3|7|11|15)$", text)] <- "critical"
  requests
}
```

The locked protocol is evaluated on the sealed test split. This access is
recorded in the gold-set ledger.

```{r}
validation <- validate_protocol(
  locked,
  gold,
  .runner = offline_coder
)

validation
```

The validation result reports holdout accuracy, its interval, category-level
performance, and parse failures. These quantities describe agreement with the
specified human labels. They do not establish that the codebook captures the
construct, and repeated holdout use appears in the gold-set ledger.

The corpus includes the audited units and new statements. The same response
function and locked protocol produce the model labels.

```{r}
new_corpus <- data.frame(
  id = sprintf("c%02d", seq_len(20)),
  text = c(
    paste("clear public benefit new", seq_len(8)),
    paste("serious public harm new", seq_len(8)),
    paste("mixed evidence new", seq_len(4))
  ),
  stringsAsFactors = FALSE
)

corpus <- rbind(gold_data[c("id", "text")], new_corpus)

coded <- code_corpus(
  corpus,
  locked,
  "text",
  id = "id",
  .runner = offline_coder
)

coded_data <- tibble::as_tibble(coded)
head(coded_data[c("id", "label")])
```

`gold_correct()` compares the model and human labels in the holdout sample.
It uses those differences to correct corpus category shares and calculate their
standard errors.

```{r}
corrected <- gold_correct(coded, gold)

tibble::as_tibble(corrected)
```

The corrected estimates use the holdout confusion matrix to adjust the corpus
shares. Their standard errors carry uncertainty from the finite holdout into
the prevalence estimates. Interpretation remains conditional on the codebook,
the human labels, and the match between holdout and corpus coding errors.

`LLMR::report()` combines the validation result with the locked protocol and
gold set. The report includes the holdout ledger, which now records the
validation and correction.

```{r}
methods <- LLMR::report(
  validation,
  gold = gold,
  protocol = locked
)

methods
```

## Robustness and replication

A robustness audit recomputes an estimand across prompt wording, label order,
and temperature. Here the estimand is the supportive share centered at 0.5.

```{r}
plan <- audit_plan(
  data = new_corpus,
  text = "text",
  estimator = function(d) {
    mean(d$label == "supportive", na.rm = TRUE) - 0.5
  },
  labels = c("supportive", "critical"),
  prompt = paste(
    "Classify as one of: {labels}.",
    "",
    "{text}",
    "",
    "Label:",
    sep = "\n"
  )
)

plan <- audit_add_models(plan, list(keyword = cfg))
plan <- audit_add_prompts(
  plan,
  concise = "Choose one label from {labels} for this statement: {text}"
)
plan <- audit_add_perturbations(
  plan,
  label_order = "reversed",
  temperature = 0.7
)

audit <- audit_run(plan, .runner = offline_coder)

audit_stability(audit)
audit_fragility(audit)
```

`audit_stability()` describes variation inside the specified grid, not across
all possible measurement choices. `audit_fragility()` measures the smallest
number of changed grid dimensions needed to alter the stated conclusion. The
fixed response function used here isolates the audit structure; a substantive
study should include defensible prompt and model alternatives.

The offline response function writes no LLMR call log, so one compact JSON
record supplies the fields needed to demonstrate archive construction.

```{r}
log_path <- tempfile(fileext = ".jsonl")
writeLines(paste0(
  '{"ts":"2026-07-01T10:00:01+0000","schema_version":"1.0","kind":"call",',
  '"provider":"groq","model":"offline-demo","status":200,',
  '"request":{"messages":[{"role":"user","content":"Classify statement 1"}],',
  '"temperature":0},"usage":{"sent":12,"rec":1},',
  '"response_id":"offline-1","text":"supportive"}'
), log_path)

archive <- archive_seal(
  archive_build(log_path, name = "offline-study")
)

archive_check(
  archive,
  results = data.frame(response_id = "offline-1")
)
```

The check compares the stored record with its content hash and seal, then
matches the result identifier to the logged call. A successful check establishes
integrity and linkage; it does not promise the same reply from a changed model
or provider.

`archive_write()` stores the records and manifest in a directory. After
`archive_read()`, `archive_check()` verifies the copy and `LLMR::report()`
summarizes it.

```{r}
archive_dir <- tempfile("offline-study-")
archive_write(archive, archive_dir)

copy <- archive_read(archive_dir)
archive_check(copy)
LLMR::report(copy)
```
