---
title: "Getting started with LLMRcontent"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with LLMRcontent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

LLMRcontent treats model labels as measured variables. A study defines the
instrument, validates it against held-out human labels, and carries the
remaining error into corpus estimates. The workflow below runs without a
provider key or a network connection.

```{r}
library(LLMRcontent)
```

## Coding with held-out validation

The codebook states the construct and its category boundaries.

```{r}
cb <- codebook(
  name = "policy appraisal",
  unit = "one statement",
  categories = list(
    cb_category("supportive", "Presents the proposal as beneficial."),
    cb_category("critical", "Presents the proposal as harmful or deficient.")
  ),
  version = "1.0"
)

cb
```

The gold set stores its split assignment and seals the test split. Stable unit
identifiers provide the later link between audited rows and the coded corpus.

```{r}
gold_data <- data.frame(
  id = sprintf("g%02d", seq_len(40)),
  text = c(
    paste("clear public benefit case", seq_len(20)),
    paste("serious public harm case", seq_len(20))
  ),
  label = rep(c("supportive", "critical"), each = 20),
  stringsAsFactors = FALSE
)

set.seed(110)
gold <- gold_set(
  gold_data,
  text = "text",
  labels = "label",
  id = "id",
  split = c(dev = 0.5, test = 0.5)
)

gold
```

A locked protocol hashes the codebook, prompt, model configuration, and parser.
The same object must be used for validation and corpus coding.

```{r}
cfg <- LLMR::llm_config("groq", "offline-demo", temperature = 0)
locked <- protocol_lock(
  protocol(cb, cfg, label = "keyword protocol")
)

locked
```

Execution functions accept a `.runner` with the data-frame contract of LLMR's
parallel runner. This deterministic runner reads the unit text carried in the
experiments and does not contact a provider.

```{r}
offline_coder <- function(experiments, ...) {
  text <- tolower(experiments$text)
  user <- vapply(experiments$messages, `[[`, "", "user")
  temperature <- vapply(experiments$config, function(x) {
    value <- x$model_params$temperature
    if (is.null(value)) 0 else as.numeric(value)
  }, numeric(1))

  label <- ifelse(
    grepl("benefit", text, fixed = TRUE),
    "supportive",
    "critical"
  )

  mixed <- grepl("mixed", text, fixed = TRUE)
  reversed <- grepl("critical, supportive", user, fixed = TRUE)
  label[mixed & xor(reversed, temperature > 0)] <- "supportive"
  label[mixed & !xor(reversed, temperature > 0)] <- "critical"

  label[grepl("benefit case (3|7|11|15)$", text)] <- "critical"

  experiments$response_text <- label
  experiments$sent_tokens <- 12L
  experiments$rec_tokens <- 1L
  experiments$success <- TRUE
  experiments
}
```

The locked protocol is evaluated on the sealed test split. This access is
recorded in the gold-set ledger.

```{r}
validation <- validate_protocol(
  locked,
  gold,
  .runner = offline_coder
)

validation
```

The corpus includes the audited units and new statements. The same runner and
locked protocol produce the model labels.

```{r}
new_corpus <- data.frame(
  id = sprintf("c%02d", seq_len(20)),
  text = c(
    paste("clear public benefit new", seq_len(8)),
    paste("serious public harm new", seq_len(8)),
    paste("mixed evidence new", seq_len(4))
  ),
  stringsAsFactors = FALSE
)

corpus <- rbind(gold_data[c("id", "text")], new_corpus)

coded <- code_corpus(
  corpus,
  locked,
  "text",
  id = "id",
  .runner = offline_coder
)

head(coded[c("id", "label")])
```

The correction combines the full set of model labels with model-gold
differences in the sealed audit sample. Its standard errors describe uncertainty
in that estimated correction.

```{r}
corrected <- gold_correct(coded, gold)

tibble::as_tibble(corrected)
```

The shared report generic reads the validation object, locked protocol, and
complete holdout ledger. The ledger now contains the validation and correction.

```{r}
methods <- LLMR::report(
  validation,
  gold = gold,
  protocol = locked
)

methods
```

## Robustness and replication

A robustness audit recomputes an estimand across prompt wording, label order,
and temperature. Here the estimand is the supportive share centered at 0.5.

```{r}
plan <- audit_plan(
  data = new_corpus,
  text = "text",
  estimator = function(d) {
    mean(d$label == "supportive", na.rm = TRUE) - 0.5
  },
  labels = c("supportive", "critical"),
  prompt = paste(
    "Classify as one of: {labels}.",
    "",
    "{text}",
    "",
    "Label:",
    sep = "\n"
  )
)

plan <- audit_add_models(plan, list(keyword = cfg))
plan <- audit_add_prompts(
  plan,
  concise = "Choose one label from {labels} for this statement: {text}"
)
plan <- audit_add_perturbations(
  plan,
  label_order = "reversed",
  temperature = 0.7
)

audit <- audit_run(plan, .runner = offline_coder)

audit_stability(audit)
audit_fragility(audit)
```

The injected runner does not create provider audit records. Two synthetic
LLMR-compatible records stand in for a live study log and exercise the archive
contract without a provider call.

```{r}
log_path <- tempfile(fileext = ".jsonl")

records <- list(
  list(
    ts = "2026-07-01T10:00:01+0000",
    schema_version = "1.0",
    kind = "call",
    provider = "groq",
    model = "offline-demo",
    status = 200L,
    request = list(
      messages = list(
        list(role = "user", content = "Classify statement 1")
      ),
      temperature = 0
    ),
    usage = list(sent = 12L, rec = 1L),
    response_id = "offline-1",
    text = "supportive"
  ),
  list(
    ts = "2026-07-01T10:00:02+0000",
    schema_version = "1.0",
    kind = "call",
    provider = "groq",
    model = "offline-demo",
    status = 200L,
    request = list(
      messages = list(
        list(role = "user", content = "Classify statement 2")
      ),
      temperature = 0
    ),
    usage = list(sent = 12L, rec = 1L),
    response_id = "offline-2",
    text = "critical"
  )
)

lines <- vapply(
  records,
  function(x) {
    as.character(jsonlite::toJSON(
      x,
      auto_unbox = TRUE,
      null = "null"
    ))
  },
  character(1)
)

writeLines(lines, log_path, useBytes = TRUE)

archive <- archive_seal(
  archive_build(log_path, name = "offline-study")
)

archive_check(
  archive,
  results = data.frame(
    response_id = c("offline-1", "offline-2")
  )
)
```

Archives are plain-text records plus a manifest. Writing and reading the
archive leaves its seal available for a fresh integrity check and report.

```{r}
archive_dir <- tempfile("offline-study-")
archive_write(archive, archive_dir)

copy <- archive_read(archive_dir)
archive_check(copy)
LLMR::report(copy)
```
